---
title: "perifix_untargeted_metabolomics"
author: "Iris Kern"
date: "2023-05-11"
output: html_document
---

#Libraries and set up
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(xcms)
library(ggplot2)
library(RaMS)
library(data.table)
library(readr)
library(rstatix)
library(RColorBrewer)
library(vegan)
library(dendextend)
library(factoextra)
library(ggalluvial)
library(ggbarribbon)
library(broom)
library(purrr)
```

# Load data and wrangle
```{r}
fls <- list.files('../../mzMLs/pos', full.names = TRUE)
fls <- as.data.frame(fls) %>% 
  filter(!str_detect(fls, "DDA"))

metadataframe <- fls %>%
  # Grab just the unique filenames and rename the column
  distinct(filename=`fls`) %>%
  # Create a new column with sample type information
  mutate(samp_type=str_extract(filename, "Blk|Smp|Std|Poo")) %>%
  # Create a new column with timepoint information (either Tfinal or T0)
  mutate(timepoint=str_extract(filename, "Tfinal|T0")) %>%
  # Extract treatment information from the filename
  # Longest one needs to go first!
  mutate(treatment=str_extract(filename, "NPF|NF|PF|NP|N|P|F|C")) %>%
  # Replace accidental "P" treaments from "Pooled" with NAs
  mutate(treatment=ifelse(samp_type=="Poo", NA, treatment)) %>% 
  mutate(colid = str_extract(filename, "Std|Poo|Blk|NPF|NF|PF|NP|N|P|F|C|Tote"))
  # mutate(colid=(levels=c("Std", "Poo", "Blk", "NPF", "NF", "PF", "NP", "N", "P", "F", "C", "Tote")))

msnexp_withpeaks <- readRDS('msnexp_withpeaks.rds')
msnexp_filled <- readRDS('msnexp_filled.rds')

set.seed(20)
```

## MS data: load in pooled samples
```{r}
# msdata <- msnexp_filled %>%
#   fileNames() %>%
#   str_subset("Poo") %>%
#   grabMSdata()
```


## Make dataframes from msnexp rds file
```{r}
peak_data_long <- msnexp_filled %>%
  chromPeaks() %>%
  as.data.frame() %>%
  rownames_to_column() %>%
  mutate(peakidx=row_number())
peak_data <- msnexp_filled %>%
  featureDefinitions() %>%
  as.data.frame() %>%
  select(mzmed, rtmed, npeaks, peakidx) %>%
  rownames_to_column("id") %>%
  unnest_longer(peakidx) %>%
  rename_with(~paste0("feat_", .x), .cols = -peakidx) %>%
  left_join(peak_data_long) %>%
  mutate(filename=basename(fileNames(msnexp_filled))[sample]) %>% 
  left_join(metadataframe %>% mutate(filename=basename(filename))) %>%
  filter(samp_type=="Smp")
```

## Filter peaks using Will's program
```{r}
good_peaks <- read_csv('chosen_feats.csv')

peak_data_filtered <- peak_data %>% 
  filter(samp_type=="Smp") %>% 
  select(compound = feat_id, filename, area = into, mz, rt, treatment) %>% 
  filter(compound%in%good_peaks$feature) %>%
  group_by(compound, filename) %>%
  mutate(n=n()) %>%
  group_by(compound) %>%
  filter(all(n==1)) %>%
  ungroup() %>%
  complete(compound, filename, fill=list(area=0)) %>%
  select(-n)
```

# NMDS Plot
## Wrangle data for NMDS
```{r}
peri_mat <- peak_data_filtered %>%
  mutate(filename=str_remove(filename, "230213_Smp_")) %>%
  group_by(compound) %>%
  mutate(norm_area=rank(area)) %>%
  pivot_wider(id_cols = filename, names_from = "compound", values_from = "norm_area") %>%
  column_to_rownames("filename") %>%
  data.matrix()
```

## Plot NMDS
```{r}
mdsout <- peri_mat %>%
  metaMDS(k = 2, autotransform = FALSE)

metadataframe <- metadataframe %>% 
  mutate(filename=str_remove(filename, "../../mzMLs/pos/230213_Smp_"))

mdsout$points %>%
  as.data.frame() %>%
  rownames_to_column("filename") %>%
  left_join(metadataframe) %>%
  mutate(col_id=case_when(
    treatment=="C"~"#dbd9db",
    treatment=="F"~"#c33225",
    treatment=="P"~"#f8dc69",
    treatment=="PF"~"#f49c14",
    treatment=="N"~"#abdfc1",
    treatment=="NP"~"#87c1ea",
    treatment=="NF"~"#29ad61",
    treatment=="NFP"~"#2a7fba", 
    TRUE~"grey50"
  )) %>%
  ggplot() +
  geom_point(aes(x=MDS1, y=MDS2, color=col_id), size=4) +
  scale_color_identity(guide = "legend") +
  theme_bw()
```

# Matching standards list
```{r}
int_stds <- read_csv('https://raw.githubusercontent.com/IngallsLabUW/Ingalls_Standards/master/Ingalls_Lab_Standards.csv') %>% 
  rename(compound = Compound_Name) %>% 
  filter(Compound_Type == "Internal Standard") %>% 
  filter(Column == "HILIC") %>% 
  mutate(Concentration_nM = Concentration_uM * 1000) %>%
  select(-c(Concentration_uM)) %>% 
  select(c(compound, mz, RT_minute, z, Concentration_nM))
```

## For loop matching internal standards to peak_data_filtered
```{r}
int_eics = list()
for (i in 1:length(int_stds$mz)) {
  mz_i <- int_stds$mz[i]
  name_i <- int_stds$compound[i]
  int_eics[[i]] <- peak_data_filtered %>%
    filter(mz%between%pmppm(mz_i, ppm = 5)) %>% 
    mutate(int_std = name_i)
}
```

## For loop matching internal standards to peak_data
```{r}
int_eics_all = list()
for (i in 1:length(int_stds$mz)) {
  mz_i <- int_stds$mz[i]
  name_i <- int_stds$compound[i]
  int_eics_all[[i]] <- peak_data %>%
    filter(mz%between%pmppm(mz_i, ppm = 5)) %>% 
    mutate(int_std = name_i)
}
```

# Metabolite Absolute and Relative Abundances
## Color setup
```{r}
nb.cols <- 10
mycolors <- colorRampPalette(brewer.pal(8, "Set1"))(nb.cols)

mycolors <- c(mycolors, "gray60")
```

## Filter for 10 most abundant compounds in peak_data_filtered and peak_data
```{r}
biggest_peaks_all <- peak_data %>% 
  arrange(desc(into)) %>%
  mutate(feat_id=ifelse(feat_id%in%unique(feat_id)[1:10], feat_id, "Other")) %>%
  mutate(feat_id=factor(feat_id, levels=unique(feat_id))) %>%
  group_by(feat_id, treatment) %>%
  summarise(into=sum(into)) %>%
  drop_na()

biggest_peaks_good <- peak_data_filtered %>%
  arrange(desc(area)) %>%
  mutate(compound=ifelse(compound%in%unique(compound)[1:10], compound, "Other")) %>%
  mutate(compound=factor(compound, levels=unique(compound))) %>%
  group_by(compound, treatment) %>%
  summarise(area=sum(area)) %>%
  drop_na()
```

## Relative metabolite abundance for peak_data
```{r}
biggest_peaks_all %>% 
  group_by(feat_id, treatment) %>%
  summarise(into=sum(into)) %>%
  ggplot() +
  geom_col(aes(x=treatment, y=into, fill = feat_id), width = 0.5, 
           position = "fill") + 
  scale_fill_manual(values = mycolors)
```

## Relative metabolite abundance for peak_data_filtered
```{r}
biggest_peaks_good %>% 
  group_by(compound, treatment) %>%
  summarise(area=sum(area)) %>%
  ggplot() +
  geom_col(aes(x=treatment, y=area, fill = compound), width = 0.5, 
           position = "fill") + 
  scale_fill_manual(values = mycolors) + 
  ggtitle("Selected Metabolites and Their Relative Abundances Across Treatments")
```

## Absolute metabolite abundace of peak_data
```{r}
biggest_peaks_all %>% 
  group_by(feat_id, treatment) %>%
  summarise(into=sum(into)) %>%
  ggplot() +
  geom_col(aes(x=treatment, y=into, fill = feat_id), width = 0.5, 
           position = "stack") + 
  scale_fill_manual(values = mycolors) +
  ggtitle("Top 10 Largest Metabolites in Unfiltered Dataset")
```

## Absolute metabolite abundance of peak_data_filtered
```{r}
biggest_peaks_good %>% 
  group_by(compound, treatment) %>%
  summarise(area=sum(area)) %>%
  ggplot() +
  geom_col(aes(x=treatment, y=area, fill = compound), width = 0.5, 
           position = "stack") + 
  scale_fill_manual(values = mycolors) +
  ggtitle("Top 10 Largest Metabolites in Filtered Good Peak Dataset")
```

# Peak areas across treatment groups for every compound in peak_data_filtered
## DON'T RUN UNLESS YOU NEED IT
```{r}
split_eics <- split(peak_data_filtered, peak_data_filtered$compound)
plot_list <- list()
for(i in 1:length(split_eics)){
  plot_list[[i]] <- ggplot(split_eics[[i]]) + 
    geom_boxplot(aes(x = treatment, y = area)) + 
    facet_wrap(~compound, scales = "free_y")
}
plot_list
```

# Hierarchical clustering
## Data wrangling for peak_data_filtered
```{r}
peak_data_numeric <- peak_data_filtered %>% 
  mutate(treatment = gsub("230213_Smp_Tfinal", "", filename)) %>%
  mutate(treatment = gsub(".mzML", "", treatment)) %>% 
  mutate(treatment = gsub("230213_Smp_T0", "", treatment)) %>% 
  mutate(compound_numeric = gsub("FT", "", compound)) %>%
  mutate(compound_numeric = as.numeric(compound_numeric)) %>%
  select("treatment", "area", "compound_numeric")

peak_data_numeric <- peak_data_numeric %>% 
  pivot_wider(names_from = treatment, values_from = area) %>% 
  remove_rownames %>% 
  column_to_rownames(var = "compound_numeric")

peak_data_scaled <- t(scale(t(peak_data_numeric)))
```

## Data wrangling for metabolites with peak area > 1e+08
```{r}
peak_data_numeric_abun <- peak_data_filtered %>% 
  mutate(treatment = gsub("230213_Smp_Tfinal", "", filename)) %>%
  mutate(treatment = gsub(".mzML", "", treatment)) %>% 
  mutate(treatment = gsub("230213_Smp_T0", "", treatment)) %>% 
  mutate(compound_numeric = gsub("FT", "", compound)) %>%
  mutate(compound_numeric = as.numeric(compound_numeric)) %>%
  select("treatment", "area", "compound_numeric") %>% 
  filter(area >= 1e+08)

peak_data_numeric_abun_wide <- peak_data_numeric_abun %>% 
  pivot_wider(names_from = treatment, values_from = area) %>% 
  remove_rownames %>% 
  column_to_rownames(var = "compound_numeric") %>% 
  mutate_all(~replace_na(.,0))

peak_data_scaled_abun <- t(scale(t(peak_data_numeric_abun_wide)))
```

## Run hierarchical clustering for all peak_data_filtered
```{r}
hclust_peak_data <- hclust(dist(peak_data_scaled, method = 'euclidean'))
```

### Dendrogram visualization of hierarchical clustering - full dataset
```{r}
plot(hclust_peak_data)

dend_hclust <- as.dendrogram(hclust_peak_data)
dend_hclust_color <- color_branches(dend_hclust, k = 8)
plot(dend_hclust_color)
```

## Run hierarchical clustering for most abundant metabolites
```{r}
hclust_peak_data_abun <- hclust(dist(peak_data_scaled_abun, method = 'euclidean'))

cut_avg <- cutree(hclust_peak_data_abun, k = 12)
```

### Dendrogram visualization of hierarchical clustering - abundant metabolites
```{r}
plot(hclust_peak_data_abun)

dend_hclust_abun <- as.dendrogram(hclust_peak_data_abun)
dend_hclust_color_abun <- color_branches(dend_hclust_abun, h = 4)
plot(dend_hclust_color_abun)
```

### Putting hierarchical clustering data into long format for additional visualization
```{r}
hclust_cl <- mutate(peak_data_numeric_abun_wide, clusters = cut_avg )

hclust_cl_long <- hclust_cl %>% 
  pivot_longer(
    cols = c("Tote_1", "Tote_2", "Tote_3", "Tote_4", "C_1", "C_2", "C_3", "N_1", "N_2",
             "N_3", "F_1", "F_2", "F_3", "NF_1", "NF_2", "NF_3", "NP_1", "NP_2", "NP_3",
             "NPF_1", "NPF_2", "NPF_3", "P_1", "P_2", "P_3", "PF_1", "PF_2", "PF_3"),
    names_to = "treatment", values_to = "area") %>% 
  filter(area != 0) %>% 
  left_join(peak_data_filtered, by = "area") %>% 
  select(-c("treatment.y", "rt", "filename")) %>% 
  rename(treatment = treatment.x)

hclust_cl_long <- hclust_cl_long %>% 
  mutate(cluster = as.character(clusters)) %>% 
  mutate(compound = str_remove(compound, "FT")) %>% 
  mutate(compound = str_remove(compound, "0")) %>% 
  mutate(compound = str_remove(compound, "00")) %>% 
  mutate(compound = str_remove(compound, "000"))
```

### Dot plot of peak areas and hierarchical clustering - abundant metabolites
```{r}
hclust_plotly <- hclust_cl_long %>% 
  group_by(treatment) %>% 
  mutate(mean_area = mean(area)) %>% 
  filter(area >= 1e+08) %>% 
  ggplot(aes(compound, treatment, size = area, color = cluster)) +
  geom_point() +
  theme_bw() + 
  ggtitle("Peak Area of Unknown Metabolites Across Varying Environmental Conditions") + 
  ylab("Added Nutrients and Replicate ID") + 
  xlab("Metabolite ID") + 
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) + 
  scale_color_manual(values = c("indianred4", "lightsalmon3", "lightsteelblue3", "skyblue", 
                                "steelblue4", "springgreen4", "olivedrab", "pink1", 
                                "plum3", "mediumpurple4", "lightgoldenrod3", "darkgoldenrod3")) + 
  theme(legend.position = "none")

ggplotly(hclust_plotly)
```

# K means clustering
## Data wrangling of peak_data_filtered
```{r}
peak_data_scaled <- t(scale(t(peak_data_numeric)))
 
peak_filtered_km <- kmeans(peak_data_scaled, centers = 8, nstart = 25)

table(peak_filtered_km$cluster)

peak_clusters <- as.data.frame(peak_filtered_km$cluster) %>% 
  rename(cluster = `peak_filtered_km$cluster`) 
```

## Data wrangling for k-means visualization
```{r}
peak_data_numeric_vis <- peak_data_numeric %>% 
  mutate(cluster = as.character(peak_filtered_km$cluster)) %>% 
  mutate(row_names = row.names(peak_data_numeric)) %>% 
  mutate(id = row_number()) %>% 
  remove_rownames %>% 
  column_to_rownames(var = "id") %>% 
  rename(compound = row_names)

peak_data_km_vis <- peak_data_numeric_vis %>% 
  pivot_longer(
    cols = c("Tote_1", "Tote_2", "Tote_3", "Tote_4", "C_1", "C_2", "C_3", "N_1", "N_2",
             "N_3", "F_1", "F_2", "F_3", "NF_1", "NF_2", "NF_3", "NP_1", "NP_2", "NP_3",
             "NPF_1", "NPF_2", "NPF_3", "P_1", "P_2", "P_3", "PF_1", "PF_2", "PF_3"),
    names_to = "treatment", values_to = "area") %>% 
  left_join(peak_data_filtered, by = "area") %>% 
  select(-c("treatment.y", "compound.y")) %>% 
  rename(treatment = treatment.x) %>% 
  rename(compound = compound.x)
```

## K-means visualization
### Visualization of different clusters in ggplot
```{r}
ggplot(subset(peak_data_km_vis, cluster %in% 8)) +
  geom_boxplot(aes(x = treatment, y = area, color = compound)) + 
  theme(legend.position = "none") 
```

### Fviz cluster
```{r}
fviz_cluster(peak_filtered_km, data = peak_data_numeric,
             geom = "point",
             ellipse.type = "convex", 
             ggtheme = theme_bw()
             )
```

### Dot plot of peak areas and k-means clustering - peak_data_filtered
```{r}
peak_data_km_vis$treatment[peak_data_km_vis$treatment == "Tote_1"] <- "T0_1"
peak_data_km_vis$treatment[peak_data_km_vis$treatment == "Tote_2"] <- "T0_2"
peak_data_km_vis$treatment[peak_data_km_vis$treatment == "Tote_3"] <- "T0_3"
peak_data_km_vis$treatment[peak_data_km_vis$treatment == "Tote_4"] <- "T0_4"

peak_data_km_vis

p <- peak_data_km_vis %>%
  group_by(compound) %>% 
  mutate(mean_area = mean(area)) %>% 
  filter(area >= 2e+08) %>% 
  ggplot(aes(compound, treatment, size = area, color = cluster)) +
  geom_point() +
  theme_bw() + 
  ggtitle("Peak Area of Unknown Metabolites Across Varying Environmental Conditions") + 
  ylab("Added Nutrients and Replicate ID") + 
  xlab("Metabolite ID") + 
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) + 
  scale_color_manual(values = c("indianred4", "lightsalmon3", "lightsteelblue3", "skyblue", 
                                "steelblue4", "springgreen4", "olivedrab", "pink1", 
                                "plum3", "mediumpurple4", "lightgoldenrod3", "darkgoldenrod3")) + 
  theme(legend.position = "none")

ggplotly(p)
```

### Dot plot of peak areas (>1e+08) and k-means clustring
```{r}
q <- peak_data_km_vis %>%
  group_by(compound) %>% 
  mutate(mean_area = mean(area)) %>% 
  filter(area >= 1e+08) %>% 
  ggplot(aes(compound, treatment, size = area, color = cluster)) +
  geom_point() +
  theme_bw() + 
  ggtitle("Peak Area of Unknown Metabolites Across Varying Environmental Conditions") + 
  ylab("Added Nutrients and Replicate ID") + 
  xlab("Metabolite ID") + 
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) + 
  scale_color_manual(values = c("indianred4", "lightsalmon3", "lightsteelblue3", "skyblue", 
                                "steelblue4", "springgreen4", "olivedrab", "pink1", 
                                "plum3", "mediumpurple4", "lightgoldenrod3", "darkgoldenrod3")) + 
  theme(legend.position = "none")

ggplotly(q)
```

# Alluvial/geom_bar_ribbon
## Data wrangle
```{r}
peak_data_vis_alluvial <- peak_data_km_vis %>% 
  select(-c("filename", "cluster", "mz", "rt")) %>% 
  mutate(treatment = str_remove(treatment, "_\\d$")) %>% 
  group_by(treatment, compound) %>% 
  mutate(area_mean_treatment = mean(area)) %>% 
  distinct(area_mean_treatment, .keep_all = TRUE) %>% 
  ungroup()

alluvial_data_wide <- peak_data_vis_alluvial %>%
  select(-c(area)) %>% 
  pivot_wider(names_from = treatment, values_from = area_mean_treatment) %>%
  remove_rownames %>% 
  # column_to_rownames(var = "compound")
  mutate_all(as.character()) %>% 
  slice(1:3)
```

## geom_bar_ribbon attempts
```{r}
peak_data_vis_alluvial %>%
  arrange(area_mean_treatment) %>%
  # filter(treatment%in%c("C", "F")) %>%
  mutate(treatment=factor(treatment, levels=unique(treatment))) %>%
  filter(compound%in%head(unique(compound), 10)) %>%
  ggplot() +
  geom_bar_ribbon(aes(x=as.numeric(treatment), y=area, fill=compound), stat = "identity", interp_res = 10) +
  scale_x_continuous(breaks = 1:9, labels = unique(peak_data_vis_alluvial$treatment)) +
  theme_bw() + 
  ylab("Relative Peak Area")
```

## Stacked bar plot
```{r}
peak_data_vis_alluvial %>% 
  arrange(area_mean_treatment) %>% 
  filter(compound%in%head(unique(compound), 10)) %>% 
  ggplot() + 
  geom_col(aes(x = treatment, y = area, fill = compound)) + 
  theme_bw()

```

# Volcano plots
## Data wrangle
```{r}
peak_data_abun_filt <- peak_data_km_vis %>% 
  group_by(compound, treatment) %>% 
  mutate(mean_area = mean(area)) %>% 
  select(-c("filename", "cluster", "mz", "rt")) %>% 
  mutate(treatment = str_remove(treatment, "_\\d$")) %>% 
  group_by(treatment, compound) %>% 
  mutate(area_mean_treatment = mean(area)) %>% 
  ungroup()
```

## Fold change calculation vs T0 for metabolites > 1e+08 peak areas
```{r}
foldchange <- data.frame(compound = character(), treatment = character(), pval = numeric(), fc = numeric())

for (compound in unique(peak_data_abun_filt$compound)) {
  for (group in unique(peak_data_abun_filt$treatment)) {
    if (group != 'T0') {
      temp <- peak_data_abun_filt[peak_data_abun_filt$compound == compound,]
      # temp <- temp[temp$treatment == group,]
      
      pval <- t.test(temp[temp$treatment == 'T0',]$area,
                     temp[temp$treatment == group,]$area)$p.value
      fc <- (log2(mean(temp[temp$treatment == group,]$area) + 1) - 
               log2(mean(temp[temp$treatment == 'T0',]$area) + 1))
      
      new_row <- data.frame(compound = compound, treatment = group, pval = pval, fc = fc)
      foldchange <- rbind(foldchange, new_row)
    }
  }
}
```

## Fold change calculation vs control for metabolites > 1e+08 peak areas
```{r}
foldchange_control <- data.frame(compound = character(), treatment = character(), pval = numeric(), fc = numeric())

for (compound in unique(peak_data_abun_filt$compound)) {
  for (group in unique(peak_data_abun_filt$treatment)) {
    if (group != 'C') {
      temp <- peak_data_abun_filt[peak_data_abun_filt$compound == compound,]
      # temp <- temp[temp$treatment == group,]
      
      pval <- t.test(temp[temp$treatment == 'C',]$area,
                     temp[temp$treatment == group,]$area)$p.value
      fc <- (log2(mean(temp[temp$treatment == group,]$area) + 1) - 
               log2(mean(temp[temp$treatment == 'C',]$area) + 1))
      
      new_row <- data.frame(compound = compound, treatment = group, pval = pval, fc = fc)
      foldchange_control <- rbind(foldchange_control, new_row)
    }
  }
}
```

## Visualizations
### Volcano plot for foldchange_control
```{r}
foldchange_control %>%
  mutate(pval = -log10(pval)) %>% 
  mutate(col_id = case_when(
    treatment == "C"~"#dbd9db",
    treatment == "F"~"#c33225",
    treatment == "P"~"#f8dc69",
    treatment == "PF"~"#f49c14",
    treatment == "N"~"#abdfc1",
    treatment == "NP"~"#87c1ea",
    treatment == "NF"~"#29ad61",
    treatment == "NPF"~"#2a7fba",
    TRUE~"gray50"
  )) %>%
  ggplot() +
  geom_point(aes(x = fc, y = pval, color = col_id)) +
  scale_color_identity() + 
  # scale_color_identity(name = "Treatment",
  #                      guide = "legend", 
  #                      labels = c("Nitrogen + Iron", 
  #                                 "Nitrogen + Phosphorus", 
  #                                 "Nitrogen", "Iron", 
  #                                 "Phosphorus + Iron", 
  #                                 "Phosphorus", "T0")) + 
  theme_bw() +
  xlim(-6, 6) + 
  xlab("Log2(Fold-Change)") + 
  ylab("-Log10(p-value)") + 
  ggtitle("fold change calculated with respect to control group")
```

### Volcano plot for foldchange
```{r}
foldchange %>%
  mutate(pval = -log10(pval)) %>% 
  mutate(col_id = case_when(
    treatment == "C"~"#dbd9db",
    treatment == "F"~"#c33225",
    treatment == "P"~"#f8dc69",
    treatment == "PF"~"#f49c14",
    treatment == "N"~"#abdfc1",
    treatment == "NP"~"#87c1ea",
    treatment == "NF"~"#29ad61",
    treatment == "NPF"~"#2a7fba"
  )) %>%
  ggplot() +
  geom_point(aes(x = fc, y = pval, color = col_id)) +
  scale_color_identity(name = "Treatment",
                       guide = "legend", 
                       labels = c("Nitrogen + Iron", "Nitrogen + Phosphorus + Iron", "Nitrogen + Phosphorus", "Nitrogen", "Iron", "Control", "Phosphorus + Iron", "Phosphorus")) + 
  theme_bw() +
  xlim(-6, 6) + 
  xlab("Log2(Fold-Change)") + 
  ylab("-Log10(p-value)") + 
  ggtitle("fold change calculated with respect to T0")
```

### Foldchange plot vs treatment
```{r}
fc_1191 <- foldchange %>% 
  group_by(treatment) %>% 
  filter(compound == "1191")

fc_1191 %>% 
  ggplot() +
  geom_point(aes(x = treatment, y = fc)) + 
  theme_bw() + 
  ggtitle("mean log2(fold change) by treatment group - fc calculated from T0") + 
  ylim(-1.5, 7) + 
  scale_x_discrete(limits = c("F", "P", "PF", "N", "NF", "NP","NPF", "C"))
```

### Foldchange_control plot cs treatment
```{r}
fc_control_1191 <- foldchange_control %>% 
  group_by(treatment) %>% 
  filter(compound == "1191")


fc_control_1191 %>% 
  ggplot() +
  geom_point(aes(x = treatment, y = fc)) + 
  theme_bw() + 
  ggtitle("mean log2(fold change) by treatment group - fc calculated from control") + 
  ylim(-1.5, 7) + 
  scale_x_discrete(limits = c("F", "P", "PF", "N", "NF", "NP","NPF", "T0"))
```

### Volcano plots faceted by treatment for foldchange
```{r}
foldchange %>%
  mutate(pval = -log10(pval)) %>% 
  ggplot() +
  geom_point(aes(x = fc, y = pval)) +
  theme_bw() + 
  facet_wrap(~treatment)
```

### Volcano plots faceted by treatment for foldchange_control
```{r}
foldchange_control %>%
  mutate(pval = -log10(pval)) %>% 
  ggplot() +
  geom_point(aes(x = fc, y = pval)) +
  theme_bw() + 
  facet_wrap(~treatment)
```

# Lots of boxplots
## Example using FT1191
```{r}
ft1191 <- peak_data %>% 
  filter(feat_id == 'FT1191')
# not in standards 

ggplot(ft1191) + 
  geom_boxplot(aes(x=treatment, y=into)) +
  geom_point(aes(x = treatment, y = into)) + 
  facet_wrap(~feat_id, scales="free_y") + 
  theme_bw()
```

## ANOVA on selected peak
```{r}
ft1191 %>% 
  anova_test(into ~ treatment)
```

# Chromatograms
## Example using FT1191
```{r}
# All peaks: biggest_peaks_all
# Good peaks: biggest_peaks_good
ft1191_mz <- 236.1493

ft1191_chr <- msnexp_filled %>% filterMz(mz=pmppm(ft1191_mz, 5)) %>% filterFile(1)
# suppressWarnings(plot(ft1191_chr))

ft1191_data <- peak_data %>% 
  filter(feat_mzmed%between%pmppm(ft1191_mz, ppm=5)) 

ft1191 <- msdata$MS1[mz%between%pmppm(ft1191_mz)] %>%
  ggplot() +
  geom_line(aes(x=rt, y=int, group=filename))

ft1191sum <- ft1191_data %>%
  group_by(feat_id) %>%
  summarize(mz=mean(feat_mzmed), rt=mean(feat_rtmed)/60)
ft1191 + 
  geom_vline(xintercept=ft1191sum$rt, color="red") + 
  ylim(c(0, 6e7)) 
```

# Isotopes
## Assign masses
```{r}
glybet <- 118.0865
glybet_iso <- 118.0865 + 1.003355
glybet_iso2 <- 118.0865 + 0.997035
```

## Glycine betaine
```{r}
peak_data %>% 
  filter(feat_mzmed%between%pmppm(glybet_iso2, ppm=5))

ft0355 <- peak_data %>% 
  filter(feat_id == "FT0355")

ft0358 <- peak_data %>% 
  filter(feat_id == "FT0358")

ftglybet<- ft0355 %>% 
  rename(into_glybet = into) %>% 
  left_join(ft0358, by = "filename") %>% 
  rename(into_glybet_iso = into) %>% 
  select(c("filename", "into_glybet", "into_glybet_iso")) %>% 
  mutate(into_ratio = into_glybet / into_glybet_iso)

# calculate error window for into_ratio. n carbons = 1/(12C/13C * nat_abund_isotope)

ggplot(ftglybet, aes(x = into_ratio)) + 
  geom_histogram() + 
  geom_vline(xintercept=20, color = "red")

glybet_data <- peak_data %>% 
  filter(feat_mzmed%between%pmppm(glybet, ppm=5)) 

glybet_ms <- msdata$MS1[mz%between%pmppm(glybet)] %>%
  ggplot() +
  geom_line(aes(x=rt, y=int, group=filename))

glybet_sum <- glybet_data %>%
  group_by(feat_id) %>%
  summarize(mz=mean(feat_mzmed), rt=mean(feat_rtmed)/60)

glybet_ms + 
  geom_vline(xintercept=glybet_sum$rt, color="red") 
```

## First glycine betaine isotope
```{r}
glybet_iso_data <- peak_data %>% 
  filter(feat_mzmed%between%pmppm(glybet_iso, ppm=5)) 

glybet_iso_ms <- msdata$MS1[mz%between%pmppm(glybet_iso)] %>%
  ggplot() +
  geom_line(aes(x=rt, y=int, group=filename))

glybet_iso_sum <- glybet_iso_data %>%
  group_by(feat_id) %>%
  summarize(mz=mean(feat_mzmed), rt=mean(feat_rtmed)/60)
glybet_iso_ms + 
  geom_vline(xintercept=glybet_iso_sum$rt, color="red") 
```

## Glycine betaine second isotope
```{r}
glybet_iso2_data <- peak_data %>% 
  filter(feat_mzmed%between%pmppm(glybet_iso2, ppm=5)) 

glybet_iso2_ms <- msdata$MS1[mz%between%pmppm(glybet_iso2)] %>%
  ggplot() +
  geom_line(aes(x=rt, y=int, group=filename))

glybet_iso2_sum <- glybet_iso2_data %>%
  group_by(feat_id) %>%
  summarize(mz=mean(feat_mzmed), rt=mean(feat_rtmed)/60)
glybet_iso2_ms + 
  geom_vline(xintercept=glybet_iso2_sum$rt, color="red") 
```

