---
title: "perifix_analysis"
author: "Iris Kern"
date: "2023-03-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Environment setup
```{r}
library(tidyverse)
library(ggplot2)
library(RaMS)
library(data.table)
library(readr)
library(rstatix)
library(RColorBrewer)
```

# Load Data
```{r}
fls <- list.files('../../mzMLs/pos', full.names = TRUE)
fls <- as.data.frame(fls) %>% 
  filter(!str_detect(fls, "DDA"))

metadataframe <- fls %>%
  # Grab just the unique filenames and rename the column
  distinct(filename=`fls`) %>%
  # Create a new column with sample type information
  mutate(samp_type=str_extract(filename, "Blk|Smp|Std|Poo")) %>%
  # Create a new column with timepoint information (either Tfinal or T0)
  mutate(timepoint=str_extract(filename, "Tfinal|T0")) %>%
  # Extract treatment information from the filename
  # Longest one needs to go first!
  mutate(treatment=str_extract(filename, "NPF|NF|PF|NP|N|P|F|C")) %>%
  # Replace accidental "P" treaments from "Pooled" with NAs
  mutate(treatment=ifelse(samp_type=="Poo", NA, treatment)) %>% 
  mutate(colid = str_extract(filename, "Std|Poo|Blk|NPF|NF|PF|NP|N|P|F|C|Tote"))
  # mutate(colid=(levels=c("Std", "Poo", "Blk", "NPF", "NF", "PF", "NP", "N", "P", "F", "C", "Tote")))

msnexp_withpeaks <- readRDS('msnexp_withpeaks.rds')
msnexp_filled <- readRDS('msnexp_filled.rds')
```

```{r}
head(rtime(msnexp_filled))
```

```{r}
mzs <- mz(msnexp_filled)

## Split the list by file
mzs_by_file <- split(mzs, f = fromFile(msnexp_filled))

length(mzs_by_file)
```

# Dataframe from msnexp file
```{r}
peak_data_long <- msnexp_filled %>%
  chromPeaks() %>%
  as.data.frame() %>%
  rownames_to_column() %>%
  mutate(peakidx=row_number())
peak_data <- msnexp_filled %>%
  featureDefinitions() %>%
  as.data.frame() %>%
  select(mzmed, rtmed, npeaks, peakidx) %>%
  rownames_to_column("id") %>%
  unnest_longer(peakidx) %>%
  rename_with(~paste0("feat_", .x), .cols = -peakidx) %>%
  left_join(peak_data_long) %>%
  mutate(filename=basename(fileNames(msnexp_filled))[sample]) %>% 
  left_join(metadataframe %>% mutate(filename=basename(filename))) %>%
  filter(samp_type=="Smp")
```

# Selecting good peaks from Will's PCA
```{r}
good_peaks <- read_csv('chosen_feats.csv')
# compound name, file name, area
```

```{r}
peak_data_filtered <- peak_data %>% 
  filter(samp_type=="Smp") %>% 
  select(compound = feat_id, filename, area = into, mz, rt, treatment) %>% 
  filter(compound%in%good_peaks$feature) %>%
  group_by(compound, filename) %>%
  mutate(n=n()) %>%
  group_by(compound) %>%
  filter(all(n==1)) %>%
  ungroup() %>%
  complete(compound, filename, fill=list(area=0)) %>%
  select(-n)
```

# Initial untargeted exploration - heatmaply, perimat, pca, mds
```{r}
library(heatmaply)
peak_data_filtered %>%
  group_by(compound) %>%
  mutate(norm_area=rank(area)) %>%
  pivot_wider(id_cols = filename, names_from = "compound", values_from = "norm_area") %>%
  column_to_rownames("filename") %>%
  data.matrix() %>%
  heatmaply()
```

```{r}
peri_mat <- peak_data_filtered %>%
  mutate(filename=str_remove(filename, "230213_Smp_")) %>%
  group_by(compound) %>%
  mutate(norm_area=rank(area)) %>%
  pivot_wider(id_cols = filename, names_from = "compound", values_from = "norm_area") %>%
  column_to_rownames("filename") %>%
  data.matrix()

# Run a PCA and plot the samples only
pcaout <- prcomp(peri_mat)
plot(pcaout$x[,1], pcaout$x[,2], type="n", xlim=c(-90, 90))
text(pcaout$x[,1], pcaout$x[,2], labels = rownames(pcaout$x))
pdf("pcaout.pdf", width=8, height = 10.5)
biplot(pcaout)
dev.off()
# color by treatment, use periscope colors
```

```{r}
library(vegan)
mdsout <- peri_mat %>%
  metaMDS(k = 2, autotransform = FALSE)

metadataframe <- metadataframe %>% 
  mutate(filename=str_remove(filename, "../../mzMLs/pos/230213_Smp_"))

mdsout$points %>%
  as.data.frame() %>%
  rownames_to_column("filename") %>%
  left_join(metadataframe) %>%
  mutate(col_id=case_when(
    treatment=="C"~"#dbd9db",
    treatment=="F"~"#c33225",
    treatment=="P"~"#f8dc69",
    treatment=="PF"~"#f49c14",
    treatment=="N"~"#abdfc1",
    treatment=="NP"~"#87c1ea",
    treatment=="NF"~"#29ad61",
    treatment=="NFP"~"#2a7fba", 
    TRUE~"grey50"
  )) %>%
  ggplot() +
  geom_point(aes(x=MDS1, y=MDS2, color=col_id), size=4) +
  scale_color_identity() +
  theme_bw()
```

```{r}
good_peak_match <- peak_data_filtered %>% 
  select(compound, mz, rt, treatment, area) %>% 
  group_by(compound) %>% 
  summarise(mz = mean(mz), rt = mean(rt))
```

# MSData
```{r}
msdata <- msnexp_filled %>%
  fileNames() %>%
  str_subset("Poo") %>%
  grabMSdata()
```

# Targeted things
```{r}
guanine_mz <- 152.057235

guanine_data <- peak_data %>% 
  filter(feat_mzmed%between%pmppm(guanine_mz, ppm=5))

ggplot(guanine_data) + 
  geom_boxplot(aes(x=treatment, y=into)) +
  facet_wrap(~feat_id, scales="free_y")
```

```{r}
gbetaine_mz <- 118.086804

gbetaine_data <- peak_data %>% 
  filter(feat_mzmed%between%pmppm(gbetaine_mz, ppm=5))

ggplot(gbetaine_data) + 
  geom_boxplot(aes(x=treatment, y=into)) +
  facet_wrap(~feat_id, scales="free_y")
```

```{r}
threonine_mz <- 120.066069

threonine_data <- peak_data %>% 
  filter(feat_mzmed%between%pmppm(threonine_mz, ppm=5))

ggplot(threonine_data) + 
  geom_boxplot(aes(x=treatment, y=into)) +
  facet_wrap(~feat_id, scales="free_y")
```

# Untargeted Pipeline
## Note: need to correct to internal standards (which correction creates the least variation) (BMIS), and subtract out blanks.
## Loading in internal standards
```{r}
dat <- peak_data_filtered

int_stds <- read_csv('https://raw.githubusercontent.com/IngallsLabUW/Ingalls_Standards/master/Ingalls_Lab_Standards.csv') %>% 
  rename(compound = Compound_Name) %>% 
  filter(Compound_Type == "Internal Standard") %>% 
  filter(Column == "HILIC") %>% 
  mutate(Concentration_nM = Concentration_uM * 1000) %>%
  select(-c(Concentration_uM)) %>% 
  select(c(compound, mz, RT_minute, z, Concentration_nM))
  
```

## BMIS for Pre-Selected Good Peaks
```{r}
int_eics = list()
for (i in 1:length(int_stds$mz)) {
  mz_i <- int_stds$mz[i]
  name_i <- int_stds$compound[i]
  int_eics[[i]] <- dat %>%
    filter(mz%between%pmppm(mz_i, ppm = 5)) %>% 
    mutate(int_std = name_i)
}
```
This finds 2 internal standards: trimethylamine n-oxide and l-arginine

## BMIS for Entire Dataset
```{r}
data_all <- peak_data %>% 
  rename("compound" = "feat_id")

int_eics_all = list()
for (i in 1:length(int_stds$mz)) {
  mz_i <- int_stds$mz[i]
  name_i <- int_stds$compound[i]
  int_eics_all[[i]] <- data_all %>%
    filter(mz%between%pmppm(mz_i, ppm = 5)) %>% 
    mutate(int_std = name_i)
}
```
This finds 9 internal standards: arsenobetaine, cytosine, dl-alanine, dl-proline, glycine betaine, guanine, homarine, trimethylamine n-oxide, l-arginine. 

# Peak abundances and statistical testing
## Filtering for "biggest" peaks for entire dataset and filtered dataset
```{r}
nb.cols <- 10
mycolors <- colorRampPalette(brewer.pal(8, "Set1"))(nb.cols)

mycolors <- c(mycolors, "gray60")

biggest_peaks_all <- peak_data %>% 
  arrange(desc(into)) %>%
  mutate(feat_id=ifelse(feat_id%in%unique(feat_id)[1:10], feat_id, "Other")) %>%
  mutate(feat_id=factor(feat_id, levels=unique(feat_id))) %>%
  group_by(feat_id, treatment) %>%
  summarise(into=sum(into)) %>%
  drop_na()

biggest_peaks_good <- peak_data_filtered %>%
  arrange(desc(area)) %>%
  mutate(compound=ifelse(compound%in%unique(compound)[1:10], compound, "Other")) %>%
  mutate(compound=factor(compound, levels=unique(compound))) %>%
  group_by(compound, treatment) %>%
  summarise(area=sum(area)) %>%
  drop_na()
```

# Abundances
## Barplots for relative/absolute abundances for entire dataset and filtered dataset
```{r}
# relative abundance
biggest_peaks_all %>% 
  group_by(feat_id, treatment) %>%
  summarise(into=sum(into)) %>%
  ggplot() +
  geom_col(aes(x=treatment, y=into, fill = feat_id), width = 0.5, 
           position = "fill") + 
  scale_fill_manual(values = mycolors)
```

```{r}
biggest_peaks_good %>% 
  group_by(compound, treatment) %>%
  summarise(area=sum(area)) %>%
  ggplot() +
  geom_col(aes(x=treatment, y=area, fill = compound), width = 0.5, 
           position = "fill") + 
  scale_fill_manual(values = mycolors) + 
  ggtitle("Selected Metabolites and Their Relative Abundances Across Treatments")
```

```{r}
# absolute abundance
biggest_peaks_all %>% 
  group_by(feat_id, treatment) %>%
  summarise(into=sum(into)) %>%
  ggplot() +
  geom_col(aes(x=treatment, y=into, fill = feat_id), width = 0.5, 
           position = "stack") + 
  scale_fill_manual(values = mycolors)
```

```{r}
biggest_peaks_good %>% 
  group_by(compound, treatment) %>%
  summarise(area=sum(area)) %>%
  ggplot() +
  geom_col(aes(x=treatment, y=area, fill = compound), width = 0.5, 
           position = "stack") + 
  scale_fill_manual(values = mycolors)
```

## Distributions for entire filtered dataset
```{r}
split_eics <- split(peak_data_filtered, peak_data_filtered$compound)
plot_list <- list()
for(i in 1:length(split_eics)){
  plot_list[[i]] <- ggplot(split_eics[[i]]) + 
    geom_boxplot(aes(x = treatment, y = area)) + 
    facet_wrap(~compound, scales = "free_y")
}
plot_list
```

## k means clustering
```{r}
# peak_data_numeric <- peak_data_filtered %>% 
#   mutate(treatment = gsub("230213_Smp_Tfinal", "", filename)) %>%
#   mutate(treatment = gsub(".mzML", "", treatment)) %>% 
#   mutate(treatment = gsub("230213_Smp_T0", "", treatment)) %>% 
#   mutate(treatment_numeric = as.numeric(as.factor(treatment))) %>% 
#   mutate(compound_numeric = gsub("FT", "", compound)) %>%
#   mutate(compound_numeric = as.numeric(compound_numeric)) %>%
#   select("treatment_numeric", "area", "compound_numeric")

peak_data_numeric <- peak_data_filtered %>% 
  mutate(treatment = gsub("230213_Smp_Tfinal", "", filename)) %>%
  mutate(treatment = gsub(".mzML", "", treatment)) %>% 
  mutate(treatment = gsub("230213_Smp_T0", "", treatment)) %>% 
  mutate(compound_numeric = gsub("FT", "", compound)) %>%
  mutate(compound_numeric = as.numeric(compound_numeric)) %>%
  select("treatment", "area", "compound_numeric")

set.seed(20)

peak_data_numeric <- peak_data_numeric %>% 
  pivot_wider(names_from = treatment, values_from = area) %>% 
  remove_rownames %>% 
  column_to_rownames(var = "compound_numeric")

peak_data_scaled <- scale(peak_data_numeric) 
 
peak_filtered_km <- kmeans(peak_data_scaled, centers = 12, nstart = 25)

table(peak_filtered_km$cluster)

peak_clusters <- as.data.frame(peak_filtered_km$cluster) %>% 
  rename(cluster = `peak_filtered_km$cluster`) 

peak_filtered_km$cluster

# write code to plot each cluster
```

```{r}
peak_data_numeric_vis <- peak_data_numeric %>% 
  mutate(cluster = as.character(peak_filtered_km$cluster)) %>% 
  mutate(row_names = row.names(peak_data_numeric)) %>% 
  mutate(id = row_number()) %>% 
  remove_rownames %>% 
  column_to_rownames(var = "id") %>% 
  rename(compound = row_names)

peak_data_km_vis <- peak_data_numeric_vis %>% 
  pivot_longer(
    cols = c("Tote_1", "Tote_2", "Tote_3", "Tote_4", "C_1", "C_2", "C_3", "N_1", "N_2",
             "N_3", "F_1", "F_2", "F_3", "NF_1", "NF_2", "NF_3", "NP_1", "NP_2", "NP_3",
             "NPF_1", "NPF_2", "NPF_3", "P_1", "P_2", "P_3", "PF_1", "PF_2", "PF_3"),
    names_to = "treatment", values_to = "area") %>% 
  left_join(peak_data_filtered, by = "area") %>% 
  select(-c("treatment.y", "compound.y")) %>% 
  rename(treatment = treatment.x) %>% 
  rename(compound = compound.x)
```

```{r}
ggplot(peak_data_km_vis) +
  geom_boxplot(aes(x = treatment, y = area, color = cluster)) 
```
# Notes
hierarchical clustering instead of k-means? create synthetic compound that behaves like how i want (i.e. low in N and P but high in NP). analyze dendrogram. group together n-added and n-deplete and do a basic t test. boxplots! upregulation of iron (iron-deplete vs iron-added), etc. Phosphorus too!

classifying softwares - identifying chemical formula of 

```{r}
ggplot(subset(peak_data_km_vis, cluster %in% 12)) +
  geom_boxplot(aes(x = treatment, y = area, color = compound)) + 
  theme(legend.position = "none") 
```




```{r}
ggplot(subset(peak_data_km_vis, cluster %in% 1)) +
  geom_col(aes(x=treatment, y=area, fill = compound), width = 0.5, 
           position = "fill")
```


## k means visualization
```{r}
library(factoextra)

fviz_cluster(peak_filtered_km, data = peak_data_numeric,
             geom = "point",
             ellipse.type = "convex", 
             ggtheme = theme_bw()
             )

# peak_data_numeric %>%
#   mutate(kmeans_class=kmeans(., centers=8)$cluster) %>%
#   plotly::plot_ly(x=~compound_numeric, y=~treatment_numeric, z=~compound_numeric, color=~factor(kmeans_class),
#                   type="scatter3d", mode="markers")
```

# Interactive Plotly
```{r}
p <- peak_data_km_vis %>%
  filter(area >= 2e+08) %>% 
  ggplot( aes(compound, treatment, size = area, color=cluster)) +
  geom_point() +
  theme_bw() + 
  ggtitle("Peak Area of Unknown Metabolites Across Varying Environmental Conditions") + 
  ylab("Added Nutrients and Replicate ID") + 
  xlab("Metabolite ID") + 
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) + 
  scale_color_manual(values = c("indianred4", "lightsalmon2", "lightsteelblue3", "skyblue", "steelblue4", "springgreen4", 
                                "olivedrab", "orchid3", "palevioletred2", "mediumpurple4", "gold", "goldenrod3"))

ggplotly(p)

```


# Boxplot distributions of most abundant compounds
## FT1191
```{r}
ft1191 <- peak_data %>% 
  filter(feat_id == 'FT1191')
# not in standards 

ggplot(ft1191) + 
  geom_boxplot(aes(x=treatment, y=into)) +
  facet_wrap(~feat_id, scales="free_y")

ft1191 %>% 
  anova_test(into ~ treatment)

# no matched standard
```

## FT0355
```{r}
# good and biggest peaks - filtering
ft0355 <- peak_data %>% 
  filter(feat_id == 'FT0355')

ggplot(ft0355) + 
  geom_boxplot(aes(x=treatment, y=into)) +
  facet_wrap(~feat_id, scales="free_y")

ft0355 %>% 
  anova_test(into ~ treatment)

# probably glycine betaine
```

## FT1245
```{r}
ft1245 <- peak_data %>% 
  filter(feat_id == 'FT1245')

ggplot(ft1245) + 
  geom_boxplot(aes(x=treatment, y=into)) +
  facet_wrap(~feat_id, scales="free_y")

ft1245 %>% 
  anova_test(into ~ treatment)

# maybe muramic acid, matches mz with a slightly shorter rt - how much would that vary?
```

## FT1310
```{r}
# good and biggest peaks - filtering
ft1310 <- peak_data %>% 
  filter(feat_id == 'FT1310')

ggplot(ft1310) + 
  geom_boxplot(aes(x=treatment, y=into)) +
  facet_wrap(~feat_id, scales="free_y")

ft1310 %>% 
  anova_test(into ~ treatment)

# no matched standard
```

## FT0772
```{r}
# good and biggest peaks - filtering
ft0772 <- peak_data %>% 
  filter(feat_id == 'FT0772')

ggplot(ft0772) + 
  geom_boxplot(aes(x=treatment, y=into)) +
  facet_wrap(~feat_id, scales="free_y")

ft0772 %>% 
  anova_test(into ~ treatment)

# carnitine has same mz but not same rt
```

## FT0538
```{r}
# good and biggest peaks - filtering
ft0538 <- peak_data %>% 
  filter(feat_id == 'FT0538')

ggplot(ft0538) + 
  geom_boxplot(aes(x=treatment, y=into)) +
  facet_wrap(~feat_id, scales="free_y")

ft0538 %>% 
  anova_test(into ~ treatment)

# beta-Alaninebetaine matches mz and rt: l-isoleucine and l-leucine both have same mz too
```

## FT0561
```{r}
# good and biggest peaks - filtering
ft0561 <- peak_data %>% 
  filter(feat_id == 'FT0561')

ggplot(ft0561) + 
  geom_point(aes(x=treatment, y=into)) +
  facet_wrap(~feat_id, scales="free_y")

ft0561 %>% 
  anova_test(into ~ treatment)

# maybe hypoxanthine, matches mz and rt
```

## FT1197
```{r}
# good and biggest peaks - filtering
ft1197 <- peak_data %>% 
  filter(feat_id == 'FT1197')

ggplot(ft1197) + 
  geom_boxplot(aes(x=treatment, y=into)) +
  facet_wrap(~feat_id, scales="free_y")

ft1197 %>% 
  anova_test(into ~ treatment)

# no matched standard
```

## FT0144
```{r}
# good and biggest peaks - filtering
ft0144 <- peak_data %>% 
  filter(feat_id == 'FT0144')

ggplot(ft0144) + 
  geom_boxplot(aes(x=treatment, y=into)) +
  facet_wrap(~feat_id, scales="free_y")

ft0144 %>% 
  anova_test(into ~ treatment)

# trimethylamine n-oxide matches mz and rt
```

## FT1020
```{r}
# good and biggest peaks - filtering
ft1020 <- peak_data %>% 
  filter(feat_id == 'FT1020')

ggplot(ft1020) + 
  geom_boxplot(aes(x=treatment, y=into)) +
  facet_wrap(~feat_id, scales="free_y")

ft1020 %>% 
  anova_test(into ~ treatment)

# Indolepyruvic acid matches mz and rt
```

# Chromatograms
## FT1191
```{r}
# All peaks: biggest_peaks_all
# Good peaks: biggest_peaks_good
ft1191_mz <- 236.1493

ft1191_chr <- msnexp_filled %>% filterMz(mz=pmppm(ft1191_mz, 5)) %>% filterFile(1)
# suppressWarnings(plot(ft1191_chr))

ft1191_data <- peak_data %>% 
  filter(feat_mzmed%between%pmppm(ft1191_mz, ppm=5)) 

ft1191 <- msdata$MS1[mz%between%pmppm(ft1191_mz)] %>%
  ggplot() +
  geom_line(aes(x=rt, y=int, group=filename))

ft1191sum <- ft1191_data %>%
  group_by(feat_id) %>%
  summarize(mz=mean(feat_mzmed), rt=mean(feat_rtmed)/60)
ft1191 + 
  geom_vline(xintercept=ft1191sum$rt, color="red") + 
  ylim(c(0, 6e7)) 
```

## FT0355
```{r}
# takes too long to make ft0355_chr, not sure why
ft0355_mz <- 118.0866

ft0355_chr <- msnexp_filled %>% filterMz(mz=pmppm(ft0355_mz, 5)) %>% filterFile(1)
# suppressWarnings(plot(ft0355_chr))

ft0355_data <- peak_data %>% 
  filter(feat_mzmed%between%pmppm(ft0355_mz, ppm=5)) 

ft0355 <- msdata$MS1[mz%between%pmppm(ft0355_mz)] %>%
  ggplot() +
  geom_line(aes(x=rt, y=int, group=filename))

ft0355sum <- ft0355_data %>%
  group_by(feat_id) %>%
  summarize(mz=mean(feat_mzmed), rt=mean(feat_rtmed)/60)
ft0355 + 
  geom_vline(xintercept=ft0355sum$rt, color="red") + 
  ylim(c(0, 6e7)) 
```

## FT1245
```{r}
ft1245_mz <- 252.1443

ft1245_chr <- msnexp_filled %>% filterMz(mz=pmppm(ft1245_mz, 5)) %>% filterFile(1)
# suppressWarnings(plot(ft1245_chr))

ft1245_data <- peak_data %>% 
  filter(feat_mzmed%between%pmppm(ft1245_mz, ppm=5)) 

ft1245 <- msdata$MS1[mz%between%pmppm(ft1245_mz)] %>%
  ggplot() +
  geom_line(aes(x=rt, y=int, group=filename))

ft1245sum <- ft1245_data %>%
  group_by(feat_id) %>%
  summarize(mz=mean(feat_mzmed), rt=mean(feat_rtmed)/60)
ft1245 + 
  geom_vline(xintercept=ft1245sum$rt, color="red") + 
  ylim(c(0, 6e7)) 
```

## FT1310
```{r}
# takes a long time, won't generate the chromatogram same error as ft0355
ft1310_mz <- 274.2742

ft1310_chr <- msnexp_filled %>% filterMz(mz=pmppm(ft1310_mz, 5)) %>% filterFile(1)
# suppressWarnings(plot(ft1310_chr))

ft1310_data <- peak_data %>% 
  filter(feat_mzmed%between%pmppm(ft1310_mz, ppm=5)) 

ft1310 <- msdata$MS1[mz%between%pmppm(ft1310_mz)] %>%
  ggplot() +
  geom_line(aes(x=rt, y=int, group=filename))

ft1310sum <- ft1310_data %>%
  group_by(feat_id) %>%
  summarize(mz=mean(feat_mzmed), rt=mean(feat_rtmed)/60)
ft1310 + 
  geom_vline(xintercept=ft1310sum$rt, color="red") + 
  ylim(c(0, 1e7)) 
```

## FT0772
```{r}
ft0772_mz <- 162.1126

ft0772_chr <- msnexp_filled %>% filterMz(mz=pmppm(ft0772_mz, 5)) %>% filterFile(1)
# suppressWarnings(plot(ft0772_chr))

ft0772_data <- peak_data %>% 
  filter(feat_mzmed%between%pmppm(ft0772_mz, ppm=5)) 

ft0772 <- msdata$MS1[mz%between%pmppm(ft0772_mz)] %>%
  ggplot() +
  geom_line(aes(x=rt, y=int, group=filename))

ft0772sum <- ft0772_data %>%
  group_by(feat_id) %>%
  summarize(mz=mean(feat_mzmed), rt=mean(feat_rtmed)/60)
ft0772 + 
  geom_vline(xintercept=ft0772sum$rt, color="red") + 
  ylim(c(0, 3e7)) 
```

## FT0538
```{r}
# same error as others
ft0538_mz <- 132.1021

ft0538_chr <- msnexp_filled %>% filterMz(mz=pmppm(ft0538_mz, 5)) %>% filterFile(1)
# suppressWarnings(plot(ft0538_chr))

ft0538_data <- peak_data %>% 
  filter(feat_mzmed%between%pmppm(ft0538_mz, ppm=5)) 

ft0538 <- msdata$MS1[mz%between%pmppm(ft0538_mz)] %>%
  ggplot() +
  geom_line(aes(x=rt, y=int, group=filename))

ft0538sum <- ft0538_data %>%
  group_by(feat_id) %>%
  summarize(mz=mean(feat_mzmed), rt=mean(feat_rtmed)/60)
ft0538 + 
  geom_vline(xintercept=ft0538sum$rt, color="red") + 
  ylim(c(0, 3e7)) 
```

## FT0561
```{r}
ft0561_mz <- 137.046

ft0561_chr <- msnexp_filled %>% filterMz(mz=pmppm(ft0561_mz, 5)) %>% filterFile(1)
# suppressWarnings(plot(ft0561_chr))

ft0561_data <- peak_data %>% 
  filter(feat_mzmed%between%pmppm(ft0561_mz, ppm=5)) 

ft0561 <- msdata$MS1[mz%between%pmppm(ft0561_mz)] %>%
  ggplot() +
  geom_line(aes(x=rt, y=int, group=filename))

ft0561sum <- ft0561_data %>%
  group_by(feat_id) %>%
  summarize(mz=mean(feat_mzmed), rt=mean(feat_rtmed)/60)
ft0561 + 
  geom_vline(xintercept=ft0561sum$rt, color="red") + 
  ylim(c(0, 1e7)) 
```

## FT1197
```{r}
ft1197_mz <- 237.1527

ft1197_chr <- msnexp_filled %>% filterMz(mz=pmppm(ft1197_mz, 5)) %>% filterFile(1)
#suppressWarnings(plot(ft1197_chr))

ft1197_data <- peak_data %>% 
  filter(feat_mzmed%between%pmppm(ft1197_mz, ppm=5)) 

ft1197 <- msdata$MS1[mz%between%pmppm(ft1197_mz)] %>%
  ggplot() +
  geom_line(aes(x=rt, y=int, group=filename))

ft1197sum <- ft1197_data %>%
  group_by(feat_id) %>%
  summarize(mz=mean(feat_mzmed), rt=mean(feat_rtmed)/60)
ft1197 + 
  geom_vline(xintercept=ft1197sum$rt, color="red") + 
  ylim(c(0, 1e7)) 
```

## FT0144
```{r}
ft0144_mz <- 85.13281

ft0144_chr <- msnexp_filled %>% filterMz(mz=pmppm(ft0144_mz, 5)) %>% filterFile(1)
#suppressWarnings(plot(ft0144_chr))

ft0144_data <- peak_data %>% 
  filter(feat_mzmed%between%pmppm(ft0144_mz, ppm=5)) 

ft0144 <- msdata$MS1[mz%between%pmppm(ft0144_mz)] %>%
  ggplot() +
  geom_line(aes(x=rt, y=int, group=filename))

ft0144sum <- ft0144_data %>%
  group_by(feat_id) %>%
  summarize(mz=mean(feat_mzmed), rt=mean(feat_rtmed)/60)
ft0144 + 
  geom_vline(xintercept=ft0144sum$rt, color="red") + 
  ylim(c(0, 4e7)) 
```

## FT1020
```{r}
ft1020_mz <- 202.1803

ft1020_chr <- msnexp_filled %>% filterMz(mz=pmppm(ft1020_mz, 5)) %>% filterFile(1)
#suppressWarnings(plot(ft1020_chr))

ft1020_data <- peak_data %>% 
  filter(feat_mzmed%between%pmppm(ft1020_mz, ppm=5)) 

ft1020 <- msdata$MS1[mz%between%pmppm(ft1020_mz)] %>%
  ggplot() +
  geom_line(aes(x=rt, y=int, group=filename))

ft1020sum <- ft1020_data %>%
  group_by(feat_id) %>%
  summarize(mz=mean(feat_mzmed), rt=mean(feat_rtmed)/60)
ft1020 + 
  geom_vline(xintercept=ft1020sum$rt, color="red") + 
  ylim(c(0, 4e7)) 
```

# Isotope/boop exploration
```{r}
boop <- 118.0865
boop2 <- 118.0865 + 1.003355
boop3 <- 118.0865 + 0.997035

peak_data %>% 
  filter(feat_mzmed%between%pmppm(boop3, ppm=5))

ft0355 <- peak_data %>% 
  filter(feat_id == "FT0355")

ft0358 <- peak_data %>% 
  filter(feat_id == "FT0358")

ftboops <- ft0355 %>% 
  rename(into_boop = into) %>% 
  left_join(ft0358, by = "filename") %>% 
  rename(into_boop2 = into) %>% 
  select(c("filename", "into_boop", "into_boop2")) %>% 
  mutate(into_ratio = into_boop / into_boop2)

# calculate error window for into_ratio. n carbons = 1/(12C/13C * nat_abund_isotope)

ggplot(ftboops, aes(x = into_ratio)) + 
  geom_histogram() + 
  geom_vline(xintercept=20, color = "red")

boop_data <- peak_data %>% 
  filter(feat_mzmed%between%pmppm(boop, ppm=5)) 

boop <- msdata$MS1[mz%between%pmppm(boop)] %>%
  ggplot() +
  geom_line(aes(x=rt, y=int, group=filename))

boopsum <- boop_data %>%
  group_by(feat_id) %>%
  summarize(mz=mean(feat_mzmed), rt=mean(feat_rtmed)/60)
boop + 
  geom_vline(xintercept=boopsum$rt, color="red") 
```

```{r}
boop2_data <- peak_data %>% 
  filter(feat_mzmed%between%pmppm(boop2, ppm=5)) 

boop2 <- msdata$MS1[mz%between%pmppm(boop2)] %>%
  ggplot() +
  geom_line(aes(x=rt, y=int, group=filename))

boop2sum <- boop2_data %>%
  group_by(feat_id) %>%
  summarize(mz=mean(feat_mzmed), rt=mean(feat_rtmed)/60)
boop2 + 
  geom_vline(xintercept=boop2sum$rt, color="red") 
```

```{r}
boop3_data <- peak_data %>% 
  filter(feat_mzmed%between%pmppm(boop3, ppm=5)) 

boop3 <- msdata$MS1[mz%between%pmppm(boop3)] %>%
  ggplot() +
  geom_line(aes(x=rt, y=int, group=filename))

boop3sum <- boop3_data %>%
  group_by(feat_id) %>%
  summarize(mz=mean(feat_mzmed), rt=mean(feat_rtmed)/60)
boop3 + 
  geom_vline(xintercept=boop3sum$rt, color="red") 
```

